Dark matter halo shape in the Auriga simulations: radial dependence, time evolution and baryon disk alignment

### 2019-07-17 Document/paper.tex
edit

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.
This effect has been already reported in other simulations \citep{Debattista08, Bryan13,
  Butsky16, Chua19, Artale19}.
The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback.
The key concept behind these results is that the baryon distribution
influences and correlates with the dark matter halo shape.

In order to explore this idea in the Auriga simulations we looked for
correlations with stellar disk properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disk to total mass ratio and
the optical radius.

Figure \ref{fig:disk_correlations} summarizes the correlations of
these quantities with the triaxility at five diferent radii.
We ues the Spearmen's rank correlation coefficient to quantify the
correlation strength.

@@

\section{Conclusions}
\section{Discussion and Conclusions}

@@

>>>>

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.
This effect has been already reported in other simulations \citep{Debattista08, Bryan13,
  Butsky16, Chua19, Artale19}.
The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback.
The key concept behind these results is that the baryon distribution
influences and correlates with the dark matter halo shape. 
The broad tendency is that massive stellar disks correlate with
spherical dark matter distributions.

In order to explore this idea in the Auriga simulations we quantify
the  correlation between halo shape with baronyc disk properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disk to total mass ratio and
the optical radius.

Figure \ref{fig:disk_correlations} summarizes the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
This confirms our expectations. The strongest correlations are found
with the halo shape measured at radii smaller than $\sim 28$kpc, which
is close to the upper limit of the disk optical radius among our
simulation sample. 
The trend is such that halos with large triaxiality correlate with
high gas density and stellar disks with low mass and small size.
In turn, massive and large stellar disks with in a low density gas
environment correlate with low triaxiliaty.

Using the six halos in the high resolution sample we find that the
memory effect put foward by \cite{VeraCiro11}, that is the correlation
between the historical and radial evolution of the shape, only holds
for a fraction of the sample and is clearly broken for two halos in
the sample.
Given the small size of the sample we use to study this effect we
cannot find any significant correlation with other disk properties in
order to understand the shape amnesia. 
This effect has not been studied in detail in other simulations. 
We consider that understanding this effect is the first promising
extension of our studies to better understand dark matter halo shape
correlation with disk properties in Milky Way type galaxies.
avenues to continue a deeper study on dark matter halo shape.

The second promising extension of our results deals with the shape
alignment as a function of radius.
In concordance

@@

\section{Conclusions}

@@

The second part of our study on the historical shape finds that in the
presence of baryons the memory effect found by \citep{VeraCiro11} does
not hold for all halos in the high resolution sample.
